\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}
\usepackage{physics}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage[margin=2.5cm]{geometry}
\usepackage{enumerate}
\usepackage{float}
\usepackage{booktabs}
\usepackage{natbib}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,shapes.geometric,fit,calc}

% Theorem environments
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{axiom}[theorem]{Axiom}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}

% Custom commands
\newcommand{\kB}{k_{\mathrm{B}}}
\newcommand{\dcat}{d_{\mathrm{cat}}}
\newcommand{\Sig}{\Sigma}
\newcommand{\Sk}{S_k}
\newcommand{\St}{S_t}
\newcommand{\Se}{S_e}
\newcommand{\Scoord}{\mathbf{S}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\morphism}{\Phi}
\newcommand{\catalyst}{\mathcal{C}}
\newcommand{\loss}{\mathcal{L}}
\newcommand{\dataset}{\mathcal{D}}
\newcommand{\model}{\mathcal{M}}
\newcommand{\params}{\theta}
\newcommand{\vocab}{\mathcal{V}}
\newcommand{\partcalc}{\textsc{PartitionCalculus}}

% Code listing style
\lstset{
    basicstyle=\ttfamily\small,
    breaklines=true,
    frame=single,
    xleftmargin=1.5em,
    framexleftmargin=1em,
    numbers=left,
    numberstyle=\tiny\color{gray},
    backgroundcolor=\color{gray!5},
    keywordstyle=\color{blue}\bfseries,
    commentstyle=\color{gray}\itshape,
    stringstyle=\color{red},
}

\title{Domain-Specific Neural Compilation for Categorical Image Access:\\
\large Trainable Morphism Chain Generation Through Knowledge Distillation in Life Science Microscopy}

\author{
Kundai Farai Sachikonye\\
\texttt{kundai.sachikonye@wzw.tum.de}
}

\date{\today}

\begin{document}

\maketitle

\begin{abstract}
We establish a mathematical framework for compiling natural language imaging specifications into executable morphism chains through domain-specific neural networks trained via knowledge distillation. The framework rests on three independently derived theoretical results: (i) the tripartite equivalence establishing that oscillation, category, and partition are isomorphic structures yielding identical entropy $S = \kB M \ln n$; (ii) the commutation relation $[\hat{O}_{\mathrm{cat}}, \hat{O}_{\mathrm{phys}}] = 0$ enabling zero-backaction categorical measurement; and (iii) the reformulation of optical image formation as discrete combinatorial constraint satisfaction over finite photon state spaces.

We prove that the mapping from imaging task specifications to morphism chains constitutes a compilation problem: the source language is constrained natural language describing biological structures; the target language is the Partition Calculus with operations \texttt{observe}, \texttt{catalyze}, \texttt{fuse}, and \texttt{access}; the compilation must preserve semantic correctness (the generated chain accesses the specified structure) and satisfy type constraints (depth compatibility, S-entropy conservation, and constraint consistency). We show that this compilation problem is learnable: the space of valid morphism chains for life science imaging is a structured subset of all possible chains, and the structure is captured by the compositional semantics of biological constraints.

The training framework employs knowledge distillation from high-capacity teacher models into compact student architectures. Teacher models generate $(t_i, \morphism_i)$ pairs where $t_i$ is a task specification and $\morphism_i$ is the corresponding morphism chain. Student models---parameter-efficient adaptations of pre-trained language models via Low-Rank Adaptation (LoRA)---learn the compilation mapping $f_\params: \mathcal{T} \to \mathcal{M}$ where $\mathcal{T}$ is the task space and $\mathcal{M}$ is the morphism chain space.

We derive the theoretical foundation for three architectural components: (1) a vision encoder bridge implementing the \texttt{observe} operation as a learned projection from image features to partition coordinates $(n, \ell, m, s)$ with capacity $C(n) = 2n^2$; (2) a catalyst selection network that reasons over biological constraints to produce optimal exclusion sequences with combined enhancement $\Delta x_{\mathrm{final}} = \Delta x_0 \cdot \prod_i \epsilon_i \cdot \exp(-\sum_{j<k} \rho_{jk})$; and (3) a cross-attention fusion mechanism whose learned attention weights recover the inter-modality correlation matrix $\rho_{ij}$, implementing the \texttt{fuse} operation.

For discrete constraint satisfaction problems arising from Refraction Puzzle Imaging, we prove that neural heuristics can reduce the effective search space from $\mathcal{O}(N_d^{T/\delta t})$ to $\mathcal{O}(\mathrm{poly}(N_d) \cdot T/\delta t)$ by learning variable ordering and constraint propagation strategies from solved instances. The amortized inference framework enables single-pass reconstruction from trained transformer networks, replacing iterative solvers.

Validation on the BBBC039 nuclei dataset demonstrates that neurally-compiled morphism chains achieve Dice coefficient 0.91, within 2\% of hand-crafted chains (0.93), while reducing specification time from expert-hours to seconds. The framework achieves 14.2$\times$ resolution enhancement through automatically selected catalyst sequences, compared to 14.87$\times$ for manually optimized configurations.

\textbf{Keywords:} morphism chain compilation, domain-specific language models, knowledge distillation, categorical imaging, partition calculus, constraint satisfaction, life science microscopy, parameter-efficient fine-tuning
\end{abstract}

\tableofcontents
\newpage

%==============================================================================
\section{Introduction}
%==============================================================================

\subsection{The Compilation Gap in Categorical Imaging}

The categorical approach to microscopy has established three foundational results. First, the tripartite equivalence---oscillation $\equiv$ category $\equiv$ partition---unifies the description of physical systems into a single mathematical framework where entropy takes the universal form $S = \kB M \ln n$ \cite{oxygen2024}. Second, the commutation relation $[\hat{O}_{\mathrm{cat}}, \hat{O}_{\mathrm{phys}}] = 0$ enables observation without physical disturbance, decoupling categorical measurement from Heisenberg uncertainty \cite{oxygen2024, multimodal2024}. Third, Refraction Puzzle Imaging recasts optical image formation as a discrete combinatorial constraint satisfaction problem, where detected intensities constrain the space of valid source configurations through finite path enumeration \cite{rpi2024}.

These results culminate in the Partition Calculus, a programming language where image analysis is not post-hoc interpretation but direct categorical structure access \cite{partition2024}. Programs are morphism chains; execution is observation. The language provides type-safe operations---\texttt{observe}, \texttt{catalyze}, \texttt{fuse}, \texttt{access}---that traverse categorical space from observed partition signatures to target biological structures.

However, a fundamental gap separates the mathematical framework from practical deployment. Constructing a morphism chain for a given imaging task requires:
\begin{enumerate}
    \item Knowledge of which biological constraints apply to the target structure
    \item Selection of appropriate catalysts with correct exclusion factors $\epsilon_i$
    \item Determination of inter-modality correlations $\rho_{ij}$ for fusion operations
    \item Verification that the composed chain satisfies depth compatibility and S-entropy conservation
\end{enumerate}

This construction currently requires expert knowledge spanning biophysics, spectroscopy, and categorical mathematics. The expert translates an imaging goal (``segment the nuclei,'' ``localize mitochondria'') into a formally correct morphism chain. This translation is, in the precise sense of compiler theory, a \emph{compilation} problem: the source language is constrained natural language describing biological structures; the target language is the Partition Calculus; the compiler must produce semantically correct, type-safe programs.

\subsection{Language Models as Compilers}

We propose that domain-specific language models, trained through knowledge distillation on the theoretical corpus of categorical imaging, can serve as compilers from natural language to morphism chains. This proposal rests on four observations:

\textbf{Observation 1: Compositionality.} Both natural language and morphism chains exhibit compositional structure. The meaning of ``segment nuclei from DAPI-stained tissue'' composes from the meanings of its parts, just as a morphism chain composes from individual operations. Language models are architecturally suited to learn compositional mappings \cite{Vaswani2017}.

\textbf{Observation 2: Structured output space.} The space of valid morphism chains is heavily constrained by the type system of Partition Calculus. Not all token sequences constitute valid chains; the type system eliminates vast regions of the output space. This structure makes the compilation problem more tractable than general-purpose code generation.

\textbf{Observation 3: Domain knowledge as parameters.} The biological constraints that govern catalyst selection---conservation laws, phase-lock continuity, thermal gradients, oxygen triangulation---are stable facts about the physical world. Rather than retrieving this knowledge at inference time (as in retrieval-augmented generation), we propose embedding it directly into model parameters through domain-specific training. This approach yields models that ``know'' the constraint landscape and can navigate it without external lookup \cite{purpose2024}.

\textbf{Observation 4: Vision-language grounding.} The \texttt{observe} operation in Partition Calculus maps images to partition signatures. Vision encoders trained on biomedical imagery learn hierarchical feature representations that naturally encode spatial structure at multiple scales. These representations can be projected into partition coordinate space, implementing \texttt{observe} as a learned function rather than a hand-crafted algorithm.

\subsection{Contributions}

This paper establishes:

\begin{enumerate}
    \item \textbf{The compilation problem} (Section~\ref{sec:compilation}): We formalize the mapping from imaging specifications to morphism chains as a compilation problem, define semantic correctness criteria, and prove that the problem is learnable under compositional assumptions.

    \item \textbf{Knowledge distillation framework} (Section~\ref{sec:distillation}): We derive a multi-stage distillation pipeline where high-capacity teacher models generate training pairs, knowledge graphs organize domain structure, and curriculum learning ensures progressive mastery from syntactic to semantic competence.

    \item \textbf{The observe bridge} (Section~\ref{sec:observe}): We prove that vision encoder features, projected through a learned linear map, satisfy the axioms required of partition signatures, establishing that neural networks can implement the \texttt{observe} operation.

    \item \textbf{Neural catalyst selection} (Section~\ref{sec:catalyst}): We show that biological constraint reasoning reduces to sequence prediction over a structured vocabulary of catalysts, and derive the training objective that ensures selected sequences achieve near-optimal resolution enhancement.

    \item \textbf{Attention as fusion} (Section~\ref{sec:fusion}): We prove that transformer cross-attention, applied to multi-modal image tokens, recovers the inter-modality correlation matrix $\rho_{ij}$, implementing the \texttt{fuse} operation with learnable correlations.

    \item \textbf{Neural-guided constraint satisfaction} (Section~\ref{sec:csp}): For the discrete CSP arising from Refraction Puzzle Imaging, we derive amortized inference architectures that replace iterative constraint propagation with single-pass neural evaluation.

    \item \textbf{Validation} (Section~\ref{sec:validation}): We demonstrate the framework on BBBC039 nuclei segmentation, comparing neurally-compiled morphism chains against hand-crafted baselines and conventional methods.
\end{enumerate}

%==============================================================================
\section{Theoretical Foundations}
\label{sec:foundations}
%==============================================================================

We present self-contained derivations of the theoretical results required for the compilation framework. These results have been established independently in prior work \cite{oxygen2024, partition2024, rpi2024, multimodal2024}; we re-derive them here in the specific form needed for neural compilation.

\subsection{Axioms}

\begin{axiom}[Bounded Phase Space]
\label{ax:bounded}
Every physical system with finite energy $E < \infty$ and finite spatial extent $V < \infty$ occupies bounded phase space with finite measure $\mu(\Gamma) < \infty$.
\end{axiom}

\begin{axiom}[Categorical Observation]
\label{ax:categorical}
An observer with finite resolution partitions phase space into a finite number of distinguishable categories. Two states belong to the same category if and only if the observer cannot distinguish them through available measurements.
\end{axiom}

These axioms are empirical facts: the first follows from the Heisenberg uncertainty principle; the second from the reproducibility of spectroscopic measurements. From these axioms, we derive the coordinate systems and measurement relations required for neural compilation.

\subsection{Partition Coordinates}

\begin{theorem}[Partition Coordinate Existence]
\label{thm:partition_coords}
Categorical partitioning of bounded spherical phase space necessarily generates coordinates $(n, \ell, m, s)$ where $n \geq 1$ is principal depth, $\ell \in \{0, \ldots, n-1\}$ is angular partition, $m \in \{-\ell, \ldots, +\ell\}$ is magnetic partition, and $s \in \{-\frac{1}{2}, +\frac{1}{2}\}$ is spin partition.
\end{theorem}

\begin{proof}
From Axiom~\ref{ax:bounded}, phase space is bounded and admits hierarchical subdivision. Spherical symmetry of the bounding region imposes nested geometric constraints. Depth $n$ measures distance from the phase space origin, requiring $n \geq 1$ for non-trivial states. Angular complexity $\ell$ cannot exceed radial depth: $\ell < n$. Orientation $m$ enumerates distinguishable angular positions within complexity level: $|m| \leq \ell$. Chirality $s$ distinguishes handedness. The capacity at depth $n$ follows from systematic counting:
\begin{equation}
C(n) = 2\sum_{\ell=0}^{n-1}(2\ell + 1) = 2n^2
\end{equation}
\end{proof}

The partition coordinates define the target representation for the \texttt{observe} bridge: a vision encoder must learn to map image features into this coordinate system.

\begin{definition}[Partition Signature]
\label{def:signature}
A partition signature $\Sig$ is a set of partition coordinates:
\begin{equation}
\Sig = \{(n_i, \ell_i, m_i, s_i)\}_{i=1}^{N}
\end{equation}
representing the categorical state of $N$ spatial positions in an image or physical system.
\end{definition}

\begin{definition}[Categorical Distance]
\label{def:cat_distance}
For states $\Sig_1 = (n_1, \ell_1, m_1, s_1)$ and $\Sig_2 = (n_2, \ell_2, m_2, s_2)$:
\begin{equation}
\dcat(\Sig_1, \Sig_2) = |n_1 - n_2| + |\ell_1 - \ell_2| + |m_1 - m_2| + |s_1 - s_2|
\end{equation}
\end{definition}

\begin{theorem}[Metric Properties]
$\dcat$ is a metric on partition coordinate space: it satisfies non-negativity, identity of indiscernibles, symmetry, and the triangle inequality.
\end{theorem}

\begin{proof}
Non-negativity and symmetry follow from absolute values. Identity of indiscernibles: $\dcat(\Sig, \Sig) = 0$ trivially, and $\dcat(\Sig_1, \Sig_2) = 0$ implies all coordinate differences vanish, hence $\Sig_1 = \Sig_2$. Triangle inequality: for any $\Sig_3$, each term $|a_1 - a_2| \leq |a_1 - a_3| + |a_3 - a_2|$, and summing yields $\dcat(\Sig_1, \Sig_2) \leq \dcat(\Sig_1, \Sig_3) + \dcat(\Sig_3, \Sig_2)$.
\end{proof}

\subsection{S-Entropy Coordinates and Conservation}

\begin{definition}[S-Entropy Coordinates]
\label{def:s_entropy}
The S-entropy coordinates $(\Sk, \St, \Se) \in [0,1]^3$ are:
\begin{align}
\Sk &= -\sum_{n,\ell,m,s} p_{n\ell ms} \log p_{n\ell ms} \quad \text{(knowledge entropy)} \\
\St &= -\sum_{\tau} p_{\tau} \log p_{\tau} \quad \text{(temporal entropy)} \\
\Se &= \sum_k \delta_k \quad \text{(evolution entropy)}
\end{align}
where $p_{n\ell ms}$ is the probability distribution over partition coordinates, $p_\tau$ is the temporal probability, and $\delta_k$ is the backaction at iteration $k$.
\end{definition}

\begin{theorem}[S-Entropy Conservation]
\label{thm:conservation}
During categorical measurement:
\begin{equation}
\Sk + \St + \Se = S_{\mathrm{total}} = \mathrm{constant}
\end{equation}
\end{theorem}

\begin{proof}
Information is neither created nor destroyed during categorical measurement, only redistributed among knowledge (spatial), temporal, and evolution components. The conservation follows from unitarity of quantum evolution and the commutation relation $[\hat{O}_{\mathrm{cat}}, \hat{O}_{\mathrm{phys}}] = 0$ (Theorem~\ref{thm:commutation}). When $\Sk$ increases (knowledge gained), $\St + \Se$ must decrease by the same amount (temporal or evolution information consumed).
\end{proof}

S-entropy conservation serves as a compile-time constraint in the neural compilation framework: any morphism chain generated by the model must preserve $S_{\mathrm{total}}$.

\subsection{Physical-Categorical Commutation}

\begin{theorem}[Commutation Relation]
\label{thm:commutation}
Categorical observables $\hat{O}_{\mathrm{cat}}$ and physical observables $\hat{O}_{\mathrm{phys}}$ commute:
\begin{equation}
[\hat{O}_{\mathrm{cat}}, \hat{O}_{\mathrm{phys}}] = 0
\end{equation}
\end{theorem}

\begin{proof}
Physical observables (position, momentum, energy) act on the quantum state $|\psi\rangle$. Categorical observables (partition coordinates, S-entropy) are functionals of the density matrix $\rho = |\psi\rangle\langle\psi|$, depending only on the eigenvalue spectrum of $\rho$, which is invariant under unitary transformations generated by physical operators. Specifically, $\hat{S}_k[\rho] = -\mathrm{Tr}(\rho \ln \rho)$ commutes with $\hat{x}$ because the von Neumann entropy depends on eigenvalues, not on the basis in which $\rho$ is expressed:
\begin{equation}
[\hat{x}, \hat{S}_k[\rho]] = \hat{x}(-\mathrm{Tr}(\rho \ln \rho)) - (-\mathrm{Tr}(\rho \ln \rho))\hat{x} = 0
\end{equation}
By Axiom~\ref{ax:categorical}, categorical measurements extract partition labels without coupling to physical degrees of freedom, confirming zero backaction.
\end{proof}

This theorem guarantees that executing a neurally-compiled morphism chain---which performs categorical observation---does not disturb the physical system under study.

\subsection{Morphism Chains and the Partition Calculus}

\begin{definition}[Morphism]
A morphism $\morphism: \Sig_A \to \Sig_B$ is a structure-preserving map between partition signatures satisfying $\morphism(\Sig_A \oplus \Sig_B) = \morphism(\Sig_A) \oplus \morphism(\Sig_B)$.
\end{definition}

\begin{definition}[Catalyst]
A catalyst $\catalyst$ is a morphism that reduces categorical distance to a target through an intermediate stage:
\begin{equation}
\dcat(\Sig_A, \Sig_B) > \dcat(\Sig_A, \catalyst(\Sig_A)) + \dcat(\catalyst(\Sig_A), \Sig_B)
\end{equation}
Each catalyst has an exclusion factor $\epsilon \in (0,1)$ quantifying configuration space reduction.
\end{definition}

\begin{definition}[Partition Calculus Operations]
\label{def:operations}
The Partition Calculus provides six typed operations:
\begin{align}
\texttt{observe} &: \mathrm{Image} \times n \to \Sig \\
\texttt{partition} &: \Sig \times n \to \Sig \\
\texttt{catalyze} &: \Sig \times \mathrm{Constraint} \to \Sig \\
\texttt{fuse} &: \Sig \times \Sig \times \rho \to \Sig \\
\texttt{access} &: \Sig \times \mathrm{Target} \to \mathrm{Structure} \\
\texttt{render} &: \Sig \times \mathrm{Modality} \to \mathrm{Image}
\end{align}
\end{definition}

Operations compose via the pipeline operator:
\begin{equation}
(f \; \texttt{|>} \; g)(x) = g(f(x))
\end{equation}

A morphism chain is a composition of operations:
\begin{equation}
\morphism_{\mathrm{total}} = \morphism_n \circ \cdots \circ \morphism_2 \circ \morphism_1
\end{equation}

\begin{theorem}[Resolution Enhancement]
\label{thm:resolution}
For $K$ catalysts with exclusion factors $\{\epsilon_i\}$ and $M$ fused modalities with pairwise correlations $\{\rho_{jk}\}$:
\begin{equation}
\Delta x_{\mathrm{final}} = \Delta x_0 \cdot \prod_{i=1}^{K} \epsilon_i \cdot \exp\left(-\sum_{j<k} \rho_{jk}\right)
\end{equation}
where $\Delta x_0$ is the diffraction-limited resolution.
\end{theorem}

\begin{proof}
Each catalyst eliminates a fraction $(1 - \epsilon_i)$ of configuration space, reducing resolution by factor $\epsilon_i$:
\begin{equation}
\Delta x_{\mathrm{after}} = \Delta x_{\mathrm{before}} \cdot \epsilon_i
\end{equation}
Sequential application yields $\prod_i \epsilon_i$. For fused modalities with correlation $\rho_{jk}$, the joint distribution narrows by $(1 - \rho_{jk}^2)^{-1/2}$. For strong correlations, $\ln(1 - \rho^2) \approx -\rho^2$, and summing over pairs with dominant linear terms gives $\exp(-\sum_{j<k} \rho_{jk})$.
\end{proof}

This theorem quantifies the objective that neural catalyst selection must optimize: choosing catalysts and correlations to minimize $\Delta x_{\mathrm{final}}$ for a given imaging target.

\subsection{Image Formation as Constraint Satisfaction}

\begin{definition}[Photon State]
At temporal index $\tau$, a photon state is:
\begin{equation}
|\psi_\tau\rangle = (\mathbf{r}_\tau, \mathbf{k}_\tau, \phi_\tau, \sigma_\tau)
\end{equation}
where $\mathbf{r}_\tau \in \mathcal{L}$ is discretized position, $\mathbf{k}_\tau$ is discretized wavevector, $\phi_\tau$ is discretized phase, and $\sigma_\tau \in \{+, -\}$ is polarization.
\end{definition}

\begin{theorem}[RPI Reconstruction]
\label{thm:rpi}
Given detected intensities $\{I(\mathbf{r}'_j)\}_{j=1}^M$, reconstructing the source distribution $S$ is equivalent to solving:
\begin{equation}
\sum_{i=1}^{N_s} A_{ji} \cdot s_i = I_j \quad \forall j
\end{equation}
where $A_{ji} = \sum_{\mathrm{paths}\; p: i \to j} w(p)$ is the path-sum transfer coefficient.
\end{theorem}

\begin{proof}
Discretization ensures $A_{ji}$ is a finite sum over the finite path space. Each path weight $w(p) = \prod_\tau T(\psi_\tau \to \psi_{\tau+1}) \cdot e^{i\Phi(p)}$ is well-defined. The inverse problem reduces to the linear system $\mathbf{A}\mathbf{s} = \mathbf{I}$.
\end{proof}

\begin{corollary}[Scattering Enhancement]
For random scattering media, the transfer matrix $\mathbf{A}$ generically achieves full rank with probability 1, guaranteeing unique reconstruction. Scattering aids rather than hinders imaging.
\end{corollary}

\subsection{Ternary State Framework for Distributed Measurement}

\begin{definition}[Ternary Molecular State]
A molecular system with two relevant electronic levels admits three categorical states:
\begin{itemize}
\item \textbf{State 0 (Absorption)}: Ground state absorbing energy
\item \textbf{State 1 (Ground)}: Thermal equilibrium
\item \textbf{State 2 (Emission)}: Excited state emitting energy
\end{itemize}
Any state admits $|\Psi\rangle = c_0|0\rangle + c_1|1\rangle + c_2|2\rangle$ with $|c_0|^2 + |c_1|^2 + |c_2|^2 = 1$.
\end{definition}

For intracellular O$_2$ with $N \sim 10^9$ molecules per cell, this provides a distributed imaging array with information content:
\begin{equation}
I = N \cdot \log_2 3 \approx 1.585 \times 10^9 \text{ bits per cycle}
\end{equation}

The ternary framework provides the measurement substrate upon which morphism chains operate. Neural compilation targets this substrate by generating chains that leverage the oxygen distribution as both detector array and computational element.

%==============================================================================
\section{The Compilation Problem}
\label{sec:compilation}
%==============================================================================

\subsection{Formal Definition}

\begin{definition}[Imaging Task Specification]
\label{def:task}
An imaging task specification $t \in \mathcal{T}$ is a tuple:
\begin{equation}
t = (\text{target}, \text{modalities}, \text{constraints}, \text{resolution})
\end{equation}
where $\text{target}$ is the biological structure to be accessed, $\text{modalities}$ is the set of available imaging modes, $\text{constraints}$ are known biological priors, and $\text{resolution}$ is the required spatial precision.
\end{definition}

In natural language, a task specification might be: ``Segment nuclei from DAPI-stained fluorescence images at sub-diffraction resolution using brightfield as auxiliary modality.'' This encodes target $=$ nuclear\_regions, modalities $= \{\text{DAPI fluorescence}, \text{brightfield}\}$, constraints $= \{\text{DNA mass conservation}, \text{chromatin phase-lock}\}$, resolution $< 200$ nm.

\begin{definition}[Morphism Chain Space]
\label{def:chain_space}
The morphism chain space $\mathcal{M}$ is the set of all well-typed compositions of Partition Calculus operations:
\begin{equation}
\mathcal{M} = \left\{ \morphism_n \circ \cdots \circ \morphism_1 \;\middle|\; \forall i: \morphism_i \in \{\texttt{observe}, \texttt{catalyze}, \texttt{fuse}, \texttt{access}\}, \; \text{type-safe} \right\}
\end{equation}
\end{definition}

\begin{definition}[Compilation Function]
A compilation function is a mapping $f: \mathcal{T} \to \mathcal{M}$ satisfying:
\begin{enumerate}
    \item \textbf{Semantic correctness}: $f(t)$ accesses the structure specified by $t.\text{target}$
    \item \textbf{Type safety}: $f(t)$ satisfies depth compatibility and constraint consistency
    \item \textbf{Conservation}: $f(t)$ preserves $\Sk + \St + \Se = S_{\mathrm{total}}$
    \item \textbf{Feasibility}: $f(t)$ uses only modalities available in $t.\text{modalities}$
\end{enumerate}
\end{definition}

\subsection{Structure of the Compilation Problem}

\begin{theorem}[Compilation Decomposition]
\label{thm:decomposition}
Any valid compilation $f(t)$ decomposes into four sequential phases:
\begin{equation}
f(t) = \texttt{access}(T) \circ \texttt{fuse}^* \circ \texttt{catalyze}^* \circ \texttt{observe}(I, n)
\end{equation}
where $\texttt{catalyze}^*$ denotes a sequence of zero or more catalyst applications and $\texttt{fuse}^*$ denotes zero or more fusion operations.
\end{theorem}

\begin{proof}
By the type system of Partition Calculus (Definition~\ref{def:operations}):
\begin{enumerate}
    \item \texttt{observe} is the unique operation mapping Image $\to$ $\Sig$; it must appear first.
    \item \texttt{catalyze} maps $\Sig \to \Sig$ and can be applied in any order (catalysts compose associatively).
    \item \texttt{fuse} maps $\Sig \times \Sig \to \Sig$ and requires signatures from potentially different modalities.
    \item \texttt{access} maps $\Sig \to$ Structure and must appear last (it terminates the chain).
\end{enumerate}
The decomposition follows from the type constraints. Within the \texttt{catalyze}$^*$ and \texttt{fuse}$^*$ phases, operations may interleave, but the overall structure is determined.
\end{proof}

This decomposition theorem is crucial for neural compilation: it reduces the generation problem from arbitrary sequence prediction to four structured sub-problems, each with constrained output spaces.

\subsection{Learnability}

\begin{theorem}[Learnability of Morphism Chain Compilation]
\label{thm:learnability}
The compilation function $f: \mathcal{T} \to \mathcal{M}$ is PAC-learnable from a polynomial number of examples under the following conditions:
\begin{enumerate}
    \item The task space $\mathcal{T}$ has finite VC dimension bounded by the number of biological constraint types
    \item The morphism chain space $\mathcal{M}$ is structured by the type system (Theorem~\ref{thm:decomposition})
    \item The compilation function is compositional: $f(t_1 \oplus t_2) = f(t_1) \circ f(t_2)$ for independent sub-tasks
\end{enumerate}
\end{theorem}

\begin{proof}
Under condition (1), $|\mathcal{T}|$ is bounded by the combinatorial product of biological structure types ($\sim 10^2$), available modalities ($\sim 10$), constraint types ($\sim 10$), and resolution levels ($\sim 10$), giving $|\mathcal{T}| \sim 10^5$.

Under condition (2), the output space decomposes per Theorem~\ref{thm:decomposition}. The \texttt{observe} phase has $\sim 10$ choices (initial depth $n$). The \texttt{catalyze}$^*$ phase selects from $\sim 10$ catalyst types with order mattering; for chain length $L$, this gives $\sim 10^L$ candidates, constrained by compatibility to $\sim L! \cdot 10$ valid orderings. The \texttt{fuse}$^*$ phase selects pairs from available modalities. The \texttt{access} phase is determined by the target.

Under condition (3), compositionality ensures that learning generalizes from component examples to novel compositions. A model that correctly compiles ``segment nuclei'' and ``detect mitochondria'' can compose them into ``segment nuclei and detect mitochondria.''

The VC dimension of the hypothesis class (parameterized morphism chains) is bounded by the product of sub-problem dimensions. By the fundamental theorem of PAC learning, a sample of size $m = \mathcal{O}(\frac{d}{\epsilon} \log \frac{d}{\epsilon})$ suffices for $\epsilon$-accuracy, where $d$ is the VC dimension.
\end{proof}

\begin{corollary}[Sample Complexity]
For $K$ catalyst types, $M$ modalities, and chain length $L \leq L_{\max}$, the number of training examples required for $\epsilon$-accurate compilation is:
\begin{equation}
m = \mathcal{O}\left(\frac{K \cdot L_{\max} + \binom{M}{2}}{\epsilon} \log \frac{K \cdot L_{\max}}{\epsilon}\right)
\end{equation}
\end{corollary}

For typical values ($K = 10$, $L_{\max} = 8$, $M = 6$, $\epsilon = 0.05$), this yields $m \sim 10^4$ training examples---well within the capacity of knowledge distillation from teacher models.

\subsection{Compilation Correctness Criteria}

\begin{definition}[Semantic Correctness]
\label{def:correctness}
A compiled chain $\morphism = f(t)$ is semantically correct if:
\begin{equation}
\morphism(\Sig_0) \cap \mathcal{S}_{\text{target}} \neq \emptyset
\end{equation}
where $\Sig_0 = \texttt{observe}(I, n)$ is the initial signature and $\mathcal{S}_{\text{target}}$ is the set of partition coordinates corresponding to the target structure.
\end{equation}
\end{definition}

\begin{definition}[Depth Adequacy]
A chain $\morphism$ is depth-adequate for target $T$ if the effective depth satisfies:
\begin{equation}
n_{\mathrm{eff}}(\morphism(\Sig_0)) \geq n_{\mathrm{req}}(T) = \frac{L_{\mathrm{field}}}{\Delta x_T}
\end{equation}
where $\Delta x_T$ is the characteristic size of $T$.
\end{definition}

\begin{definition}[Conservation Compliance]
A chain $\morphism$ is conservation-compliant if at every intermediate stage $k$:
\begin{equation}
|\Sk^{(k)} + \St^{(k)} + \Se^{(k)} - S_{\mathrm{total}}| < \epsilon_{\mathrm{machine}}
\end{equation}
\end{definition}

These criteria define the loss function for training the neural compiler (Section~\ref{sec:distillation}).

%==============================================================================
\section{Knowledge Distillation Framework}
\label{sec:distillation}
%==============================================================================

\subsection{Teacher-Student Architecture}

The knowledge distillation framework transfers domain expertise from high-capacity teacher models to compact, deployable student models. This transfer occurs through the medium of training data: teachers generate high-quality $(t_i, \morphism_i)$ pairs that students learn to reproduce.

\begin{definition}[Teacher Model]
\label{def:teacher}
A teacher model $\model_T$ with parameters $\params_T$ is a high-capacity language model capable of:
\begin{enumerate}
    \item Parsing imaging task specifications in natural language
    \item Reasoning about biological constraints and their exclusion factors
    \item Generating syntactically and semantically correct Partition Calculus programs
    \item Verifying conservation compliance and depth adequacy
\end{enumerate}
\end{definition}

Teacher models operate through in-context learning: the theoretical corpus of categorical imaging (the papers establishing partition coordinates, S-entropy conservation, catalyst mechanisms, and constraint satisfaction) is provided as context. The teacher generates morphism chains by reasoning over this context.

\begin{definition}[Student Model]
\label{def:student}
A student model $\model_S$ with parameters $\params_S = \params_0 + \Delta\params$ is a parameter-efficient adaptation of a pre-trained base model, where $\params_0$ are frozen pre-trained weights and $\Delta\params$ are trainable low-rank perturbations.
\end{definition}

\subsection{Low-Rank Adaptation}

Parameter-efficient fine-tuning through Low-Rank Adaptation (LoRA) modifies weight matrices $W \in \R^{d \times k}$ with low-rank updates \cite{Hu2022}:

\begin{equation}
W' = W + \Delta W = W + BA
\end{equation}

where $B \in \R^{d \times r}$, $A \in \R^{r \times k}$, and $r \ll \min(d, k)$ is the adaptation rank.

\begin{theorem}[LoRA Expressiveness for Morphism Chain Generation]
\label{thm:lora}
For a pre-trained language model with hidden dimension $d$ and a target compilation function with $K$ distinct catalyst types, $M$ modalities, and maximum chain length $L$, a LoRA adaptation with rank:
\begin{equation}
r \geq K + M + L + \lceil \log_2 C(n_{\max}) \rceil
\end{equation}
is sufficient to represent the compilation mapping.
\end{theorem}

\begin{proof}
The compilation mapping requires encoding:
\begin{itemize}
    \item $K$ binary decisions (include/exclude each catalyst type): $K$ dimensions
    \item $M$ continuous correlation values $\rho_{ij}$: at most $\binom{M}{2} \leq M$ independent values
    \item $L$ ordinal positions in the chain: $L$ dimensions
    \item Partition depth selection: $\lceil \log_2 C(n_{\max}) \rceil$ bits
\end{itemize}

A rank-$r$ update to the attention weight matrices provides $r \cdot d$ additional parameters, which is sufficient to linearly embed the compilation space (dimension $K + M + L + \log_2 C(n_{\max})$) into the model's hidden space (dimension $d$) provided $r$ exceeds the compilation space dimension.
\end{proof}

For typical values ($K = 10$, $M = 6$, $L = 8$, $n_{\max} = 100$, $C(100) = 20000$), we obtain $r \geq 39$. In practice, $r = 64$ provides ample capacity with computational efficiency.

\subsection{Training Data Generation}

\begin{definition}[Distillation Corpus]
\label{def:corpus}
The distillation corpus $\dataset$ consists of pairs:
\begin{equation}
\dataset = \{(t_i, \morphism_i)\}_{i=1}^{N}
\end{equation}
where $t_i$ is a natural language task specification and $\morphism_i$ is the corresponding verified morphism chain.
\end{definition}

Teacher models generate the corpus through a structured protocol:

\begin{algorithm}[H]
\caption{Distillation Corpus Generation}
\label{alg:corpus}
\begin{algorithmic}[1]
\REQUIRE Teacher model $\model_T$, theoretical corpus $\mathcal{C}$, biological structure catalog $\mathcal{B}$
\ENSURE Distillation corpus $\dataset$
\STATE $\dataset \gets \emptyset$
\FOR{each structure $b \in \mathcal{B}$}
    \FOR{each modality combination $M \subseteq \mathcal{M}_{\text{available}}$}
        \STATE $t \gets \text{GenerateTaskSpec}(b, M)$ \COMMENT{Natural language specification}
        \STATE $\morphism \gets \model_T(\mathcal{C}, t)$ \COMMENT{Teacher generates chain}
        \IF{$\text{TypeCheck}(\morphism)$ \AND $\text{ConservationCheck}(\morphism)$ \AND $\text{DepthCheck}(\morphism, b)$}
            \STATE $\dataset \gets \dataset \cup \{(t, \morphism)\}$
        \ENDIF
    \ENDFOR
\ENDFOR
\RETURN $\dataset$
\end{algorithmic}
\end{algorithm}

The verification steps (TypeCheck, ConservationCheck, DepthCheck) ensure that only correct compilations enter the training set, preventing the student from learning invalid chains.

\subsection{Knowledge Graph Construction}

Beyond flat $(t, \morphism)$ pairs, domain knowledge has relational structure that aids learning. We construct a knowledge graph encoding the relationships between biological structures, constraints, and catalysts.

\begin{definition}[Domain Knowledge Graph]
\label{def:kg}
The domain knowledge graph $\mathcal{G} = (V, E, R)$ has:
\begin{itemize}
    \item Vertices $V = V_{\text{struct}} \cup V_{\text{catalyst}} \cup V_{\text{modality}} \cup V_{\text{constraint}}$
    \item Edges $E \subseteq V \times V$ connecting related entities
    \item Relations $R$: \texttt{requires}, \texttt{enhances}, \texttt{excludes}, \texttt{correlates}, \texttt{composes}
\end{itemize}
\end{definition}

\begin{example}[Knowledge Graph Fragment]
\begin{align}
&(\text{nucleus}, \texttt{requires}, \text{chromatin\_phase\_lock}) \\
&(\text{chromatin\_phase\_lock}, \texttt{enhances}, \text{DNA\_conservation}) \\
&(\text{mitochondria}, \texttt{requires}, \text{thermal\_metabolic}) \\
&(\text{thermal\_metabolic}, \texttt{correlates}, \text{oxygen\_consumption}, \rho = 0.7) \\
&(\text{mass\_conservation}, \texttt{composes}, \text{charge\_conservation})
\end{align}
\end{example}

The knowledge graph serves two purposes: (1) it constrains the teacher's generation to biologically plausible catalyst sequences, and (2) it provides structured training signal for the student through graph-aware training objectives.

\subsection{Curriculum Learning}

Domain-specific training proceeds through a curriculum of increasing complexity \cite{Bengio2009}:

\begin{definition}[Training Curriculum]
\label{def:curriculum}
The curriculum $\mathcal{Q} = (\mathcal{Q}_1, \mathcal{Q}_2, \mathcal{Q}_3, \mathcal{Q}_4)$ comprises four stages:
\begin{enumerate}
    \item \textbf{Syntactic stage} $\mathcal{Q}_1$: Learn Partition Calculus syntax (operation names, argument types, pipeline composition)
    \item \textbf{Single-catalyst stage} $\mathcal{Q}_2$: Learn individual catalyst semantics and exclusion factors
    \item \textbf{Multi-catalyst stage} $\mathcal{Q}_3$: Learn catalyst composition, ordering effects, and constraint compatibility
    \item \textbf{Full compilation stage} $\mathcal{Q}_4$: Learn end-to-end task-to-chain mapping with fusion and resolution optimization
\end{enumerate}
\end{definition}

\begin{theorem}[Curriculum Convergence]
\label{thm:curriculum}
Under curriculum learning with stages $\mathcal{Q}_1 \to \mathcal{Q}_2 \to \mathcal{Q}_3 \to \mathcal{Q}_4$, the student model converges to $\epsilon$-optimal compilation with sample complexity:
\begin{equation}
m_{\text{curriculum}} = \sum_{i=1}^{4} m_i \leq \frac{1}{4} m_{\text{uniform}}
\end{equation}
where $m_{\text{uniform}}$ is the sample complexity for uniform (non-curriculum) training.
\end{theorem}

\begin{proof}
Each curriculum stage restricts the hypothesis class to a subset of the full compilation space. Stage $\mathcal{Q}_1$ restricts to syntactic well-formedness (finite-state grammar); $\mathcal{Q}_2$ restricts to single-operation semantics; $\mathcal{Q}_3$ restricts to bounded-length compositions; $\mathcal{Q}_4$ operates on the full space but with warm-started parameters.

The VC dimension at stage $i$ satisfies $d_i \leq d_4 / 2^{4-i}$ because each stage doubles the hypothesis space. By PAC learning theory, $m_i = \mathcal{O}(d_i / \epsilon_i)$. With the warm-start advantage, $\epsilon_i$ at stage $i$ is reduced by the progress from stage $i-1$, yielding geometric convergence. Summing a geometric series gives $\sum m_i \leq 2m_4 \leq m_{\text{uniform}} / 4$.
\end{proof}

\subsection{Training Objective}

The student model is trained to minimize a composite loss:

\begin{equation}
\loss(\params_S) = \loss_{\text{gen}} + \lambda_1 \loss_{\text{type}} + \lambda_2 \loss_{\text{cons}} + \lambda_3 \loss_{\text{res}}
\end{equation}

where:
\begin{align}
\loss_{\text{gen}} &= -\sum_{i=1}^{N} \sum_{j=1}^{|\morphism_i|} \log p_{\params_S}(\morphism_i^{(j)} | t_i, \morphism_i^{(<j)}) \label{eq:gen_loss} \\
\loss_{\text{type}} &= \sum_{i=1}^{N} \mathbb{1}[\text{TypeCheck}(f_{\params_S}(t_i)) = \text{FAIL}] \label{eq:type_loss} \\
\loss_{\text{cons}} &= \sum_{i=1}^{N} |\Sk + \St + \Se - S_{\text{total}}|^2 \label{eq:cons_loss} \\
\loss_{\text{res}} &= \sum_{i=1}^{N} \max(0, \Delta x_{\text{achieved}} - \Delta x_{\text{required}})^2 \label{eq:res_loss}
\end{align}

The generation loss $\loss_{\text{gen}}$ (Eq.~\ref{eq:gen_loss}) is the standard autoregressive cross-entropy, ensuring the student reproduces the teacher's morphism chains token by token. The type loss $\loss_{\text{type}}$ (Eq.~\ref{eq:type_loss}) penalizes chains that fail type checking. The conservation loss $\loss_{\text{cons}}$ (Eq.~\ref{eq:cons_loss}) enforces S-entropy conservation. The resolution loss $\loss_{\text{res}}$ (Eq.~\ref{eq:res_loss}) penalizes insufficient resolution enhancement.

\begin{theorem}[Loss Function Completeness]
\label{thm:loss_complete}
A model achieving $\loss(\params_S) = 0$ produces compilations satisfying all correctness criteria of Definition~\ref{def:correctness}.
\end{theorem}

\begin{proof}
$\loss_{\text{gen}} = 0$ implies the model reproduces teacher outputs exactly, which are verified correct by Algorithm~\ref{alg:corpus}. $\loss_{\text{type}} = 0$ implies all generated chains pass type checking. $\loss_{\text{cons}} = 0$ implies S-entropy conservation. $\loss_{\text{res}} = 0$ implies sufficient resolution for the target. Together, these satisfy semantic correctness, type safety, conservation compliance, and feasibility.
\end{proof}

%==============================================================================
\section{The Observe Bridge: Vision Encoders as Partition Signature Extractors}
\label{sec:observe}
%==============================================================================

\subsection{From Image Features to Partition Coordinates}

The \texttt{observe} operation maps an image $I$ to a partition signature $\Sig$ at depth $n$. We show that this mapping can be implemented by a vision encoder followed by a learned projection.

\begin{definition}[Vision Encoder]
\label{def:encoder}
A vision encoder $E: \R^{H \times W \times C} \to \R^{P \times d}$ maps an image of height $H$, width $W$, and $C$ channels to a sequence of $P$ patch embeddings, each of dimension $d$. We consider encoders pre-trained through self-supervised objectives on biomedical imagery.
\end{definition}

\begin{definition}[Partition Projection]
\label{def:projection}
A partition projection $\Pi: \R^d \to \Z^+ \times \Z_{\geq 0} \times \Z \times \{-\frac{1}{2}, +\frac{1}{2}\}$ maps each patch embedding to partition coordinates:
\begin{equation}
\Pi(\mathbf{z}) = \left(\lfloor \sigma(w_n^\top \mathbf{z} + b_n) \cdot n_{\max} \rfloor + 1, \; f_\ell(\mathbf{z}), \; f_m(\mathbf{z}), \; \text{sign}(w_s^\top \mathbf{z})\right)
\end{equation}
where $\sigma$ is the sigmoid function, $w_n, w_s \in \R^d$ and $b_n \in \R$ are learned parameters, and $f_\ell, f_m$ are constraint-satisfying mappings ensuring $0 \leq \ell < n$ and $|m| \leq \ell$.
\end{definition}

\begin{theorem}[Observe Bridge Validity]
\label{thm:observe}
The composition $\texttt{observe}_\theta = \Pi \circ E$ satisfies the axioms required of a partition signature extraction function:
\begin{enumerate}
    \item \textbf{Partition validity}: Output coordinates satisfy $n \geq 1$, $0 \leq \ell < n$, $|m| \leq \ell$, $s \in \{-\frac{1}{2}, +\frac{1}{2}\}$
    \item \textbf{Capacity consistency}: The number of distinct signatures at depth $n$ does not exceed $C(n) = 2n^2$
    \item \textbf{Locality}: Spatially proximate image patches map to categorically proximate signatures
    \item \textbf{Equivariance}: Rotation of the input image induces predictable changes in $(\ell, m)$ coordinates
\end{enumerate}
\end{theorem}

\begin{proof}
(1) The floor function with offset ensures $n \geq 1$. The constraint functions $f_\ell$ and $f_m$ are defined to satisfy range constraints by construction:
\begin{align}
f_\ell(\mathbf{z}) &= \lfloor \sigma(w_\ell^\top \mathbf{z}) \cdot (n(\mathbf{z}) - 1) \rfloor \\
f_m(\mathbf{z}) &= \lfloor \sigma(w_m^\top \mathbf{z}) \cdot (2\ell(\mathbf{z}) + 1) \rfloor - \ell(\mathbf{z})
\end{align}

(2) For depth $n$, the total number of valid $(l, m, s)$ combinations is $\sum_{\ell=0}^{n-1} (2\ell + 1) \cdot 2 = 2n^2 = C(n)$. The projection cannot produce more distinct signatures than this.

(3) Vision encoders trained with contrastive objectives (e.g., DINOv2 \cite{Oquab2024}) learn features where spatial proximity correlates with feature similarity. Formally, for patches at positions $\mathbf{r}_1, \mathbf{r}_2$ with $|\mathbf{r}_1 - \mathbf{r}_2| < \delta$:
\begin{equation}
\|E(\mathbf{r}_1) - E(\mathbf{r}_2)\| \leq L \cdot |\mathbf{r}_1 - \mathbf{r}_2|
\end{equation}
for Lipschitz constant $L$. A Lipschitz-continuous projection $\Pi$ then ensures:
\begin{equation}
\dcat(\Pi(E(\mathbf{r}_1)), \Pi(E(\mathbf{r}_2))) \leq L' \cdot |\mathbf{r}_1 - \mathbf{r}_2|
\end{equation}

(4) Vision Transformers with rotary position embeddings encode angular information in the phase of their representations. Rotation by angle $\alpha$ transforms features as $E(R_\alpha I) = U_\alpha E(I)$ for unitary $U_\alpha$. The angular partition coordinates $(\ell, m)$ are designed to encode angular information, so the projection $\Pi$ maps the unitary action to coordinate changes in $(\ell, m)$.
\end{proof}

\subsection{Hierarchical Feature Correspondence}

Self-supervised vision encoders learn hierarchical representations across transformer layers. We establish a correspondence between this hierarchy and partition depth.

\begin{theorem}[Layer-Depth Correspondence]
\label{thm:layer_depth}
For a vision transformer with $L$ layers, features at layer $l$ correspond to partition depth:
\begin{equation}
n(l) = n_{\max} \cdot \frac{l}{L}
\end{equation}
In particular, early layers encode coarse spatial structure (low $n$) and late layers encode fine structure (high $n$), consistent with the capacity relation $C(n) = 2n^2$.
\end{theorem}

\begin{proof}
Vision transformers progressively refine representations: layer $l$ captures structure at spatial scale $\sim \lambda / l$ where $\lambda$ is the receptive field. The number of distinguishable features at layer $l$ scales as $l^2$ (quadratic growth from attention composition), matching the capacity relation $C(n) = 2n^2$ with $n \propto l$. The correspondence $n(l) = n_{\max} \cdot l/L$ ensures that the full depth range is covered by the network, and the resolution at each depth is:
\begin{equation}
\Delta x(l) = \frac{L_{\text{field}}}{C(n(l))} = \frac{L_{\text{field}}}{2n_{\max}^2 l^2 / L^2}
\end{equation}
which decreases quadratically with layer index, consistent with progressive refinement.
\end{proof}

\begin{corollary}[Multi-Scale Observe]
By tapping features at multiple layers, the observe bridge can extract partition signatures at multiple depths simultaneously:
\begin{equation}
\Sig_{\text{multi}} = \bigcup_{l=1}^{L} \Pi_l(E_l(I))
\end{equation}
providing a multi-resolution partition signature without explicit multi-scale image processing.
\end{corollary}

\subsection{Training the Observe Bridge}

The partition projection $\Pi$ is trained jointly with the student model using a reconstruction objective:

\begin{equation}
\loss_{\text{observe}} = \sum_{i=1}^{N} \dcat(\Pi(E(I_i)), \Sig_i^{\text{GT}})^2
\end{equation}

where $\Sig_i^{\text{GT}}$ is the ground-truth partition signature derived from expert annotation or from the theoretical framework (e.g., known partition coordinates of nuclear boundaries, membrane structures, etc.).

For unlabeled images, we use a self-supervised objective based on the capacity constraint:

\begin{equation}
\loss_{\text{capacity}} = \sum_n \left(|\{i : n_i = n\}| - C(n)\right)^2
\end{equation}

enforcing that the distribution of partition depths in the output matches the theoretical capacity $C(n) = 2n^2$.

%==============================================================================
\section{Neural Catalyst Selection}
\label{sec:catalyst}
%==============================================================================

\subsection{The Catalyst Selection Problem}

Given a target structure $T$ and initial signature $\Sig_0$, the catalyst selection problem is to find the sequence of catalysts $(\catalyst_1, \ldots, \catalyst_K)$ that minimizes the final resolution $\Delta x_{\mathrm{final}}$ while satisfying constraint compatibility.

\begin{definition}[Catalyst Vocabulary]
\label{def:catalyst_vocab}
The catalyst vocabulary $\vocab_\catalyst$ is the set of available catalysts with their exclusion factors:
\begin{equation}
\vocab_\catalyst = \{(\catalyst_i, \epsilon_i, \text{type}_i, \text{compat}_i)\}_{i=1}^{|\vocab_\catalyst|}
\end{equation}
where $\text{type}_i \in \{\text{conservation}, \text{phase-lock}, \text{thermal}, \text{temporal}, \text{oxygen}\}$ and $\text{compat}_i \subseteq \vocab_\catalyst$ specifies compatible catalysts.
\end{definition}

\begin{table}[H]
\centering
\caption{Catalyst vocabulary for life science microscopy}
\label{tab:catalysts}
\begin{tabular}{@{}llll@{}}
\toprule
Catalyst & Type & $\epsilon$ & Biological Basis \\
\midrule
\texttt{conservation(mass)} & Conservation & 0.25 & Mass continuity \\
\texttt{conservation(charge)} & Conservation & 0.20 & Electroneutrality \\
\texttt{conservation(energy)} & Conservation & 0.10 & ATP/metabolic balance \\
\texttt{phase\_lock(membrane)} & Phase-lock & 0.15 & Lipid bilayer topology \\
\texttt{phase\_lock(chromatin)} & Phase-lock & 0.15 & DNA organization \\
\texttt{phase\_lock(cytoskeleton)} & Phase-lock & 0.20 & Actin/tubulin networks \\
\texttt{thermal(metabolic)} & Thermal & 0.10 & Heat from ATP hydrolysis \\
\texttt{thermal(gradient)} & Thermal & 0.15 & Temperature distribution \\
\texttt{temporal(cell\_cycle)} & Temporal & 0.10 & Division timing \\
\texttt{temporal(signaling)} & Temporal & 0.15 & Cascade dynamics \\
\texttt{o2(triangulation)} & Oxygen & 0.05 & Molecular O$_2$ positioning \\
\texttt{o2(consumption)} & Oxygen & 0.10 & Metabolic uptake \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Optimal Catalyst Ordering}

\begin{theorem}[Catalyst Order Independence for Resolution]
\label{thm:order_independence}
The final resolution $\Delta x_{\mathrm{final}}$ is independent of catalyst application order:
\begin{equation}
\Delta x_{\mathrm{final}} = \Delta x_0 \cdot \prod_{i=1}^{K} \epsilon_i
\end{equation}
regardless of permutation $\sigma$ applied to the catalyst sequence.
\end{theorem}

\begin{proof}
The product $\prod_i \epsilon_i$ is commutative and associative. Each catalyst reduces the configuration space independently by factor $\epsilon_i$. The order does not affect the final product.
\end{proof}

\begin{remark}
While resolution is order-independent, \emph{computational efficiency} is not. Applying the strongest catalyst (smallest $\epsilon$) first maximally reduces the search space for subsequent operations. The neural compiler should learn to emit efficient orderings even though correctness does not depend on order.
\end{remark}

\begin{theorem}[Optimal Catalyst Subset Selection]
\label{thm:optimal_subset}
Given target structure $T$ with required resolution $\Delta x_T$, the minimum catalyst set $\mathcal{C}^*$ satisfying $\Delta x_{\mathrm{final}} \leq \Delta x_T$ is the solution to:
\begin{equation}
\mathcal{C}^* = \argmin_{S \subseteq \vocab_\catalyst} |S| \quad \text{s.t.} \quad \Delta x_0 \cdot \prod_{c \in S} \epsilon_c \leq \Delta x_T \quad \text{and} \quad S \text{ is compatible}
\end{equation}
\end{theorem}

\begin{proof}
Taking logarithms, the constraint becomes $\log \Delta x_0 + \sum_{c \in S} \log \epsilon_c \leq \log \Delta x_T$. Since $\epsilon_c < 1$, $\log \epsilon_c < 0$, and this is a covering problem: find the minimum cardinality subset whose $\log \epsilon_c$ values sum to at most $\log(\Delta x_T / \Delta x_0)$. With the compatibility constraint, this reduces to minimum set cover on the compatibility graph, which is NP-hard in general but tractable for the small catalyst vocabularies ($|\vocab_\catalyst| \sim 12$) encountered in practice.
\end{proof}

\subsection{Neural Catalyst Sequence Generation}

The catalyst selection network is an autoregressive decoder that generates catalyst tokens from the catalyst vocabulary, conditioned on the task specification and current partition signature.

\begin{definition}[Catalyst Decoder]
\label{def:catalyst_decoder}
The catalyst decoder $D_\catalyst: \mathcal{T} \times \Sig \to \vocab_\catalyst^*$ generates a sequence of catalyst tokens:
\begin{equation}
p(\catalyst_1, \ldots, \catalyst_K | t, \Sig_0) = \prod_{k=1}^{K} p_\params(\catalyst_k | t, \Sig_0, \catalyst_1, \ldots, \catalyst_{k-1})
\end{equation}
with a special \texttt{<STOP>} token indicating sequence termination.
\end{definition}

The decoder is trained with the generation loss (Eq.~\ref{eq:gen_loss}) restricted to catalyst tokens, plus a resolution reward:

\begin{equation}
\loss_{\text{catalyst}} = \loss_{\text{gen}}|_{\text{catalysts}} - \lambda_r \cdot \log\left(\frac{\Delta x_0}{\Delta x_{\mathrm{final}}}\right)
\end{equation}

The second term provides signal proportional to the log resolution enhancement, encouraging the model to select maximally constraining catalyst sequences.

\subsection{Constraint Compatibility Enforcement}

The compatibility constraints from Table~\ref{tab:catalysts} define a graph $G_{\text{compat}} = (\vocab_\catalyst, E_{\text{compat}})$ where edges connect compatible catalyst pairs. The decoder must generate sequences that form cliques in this graph.

\begin{theorem}[Constrained Decoding]
\label{thm:constrained_decoding}
At each generation step $k$, the valid next-token set is:
\begin{equation}
\vocab_\catalyst^{(k)} = \{\catalyst \in \vocab_\catalyst : \catalyst \text{ compatible with } \catalyst_1, \ldots, \catalyst_{k-1}\}
\end{equation}
The model's output distribution is masked to $\vocab_\catalyst^{(k)}$ before sampling:
\begin{equation}
p'(\catalyst_k | \cdot) = \frac{p(\catalyst_k | \cdot) \cdot \mathbb{1}[\catalyst_k \in \vocab_\catalyst^{(k)}]}{\sum_{\catalyst \in \vocab_\catalyst^{(k)}} p(\catalyst | \cdot)}
\end{equation}
\end{theorem}

This constrained decoding guarantees that generated catalyst sequences are always compatible, regardless of model quality. The type system of Partition Calculus is enforced at inference time, not merely learned from data.

%==============================================================================
\section{Transformer Attention as the Fuse Operation}
\label{sec:fusion}
%==============================================================================

\subsection{Multi-Modal Partition Signature Fusion}

The \texttt{fuse} operation combines partition signatures from multiple imaging modalities, weighted by inter-modality correlations $\rho_{ij}$. We show that transformer cross-attention implements this operation with learnable correlations.

\begin{definition}[Multi-Modal Token Sequence]
Given $M$ imaging modalities, each producing $P$ patch embeddings of dimension $d$, the multi-modal token sequence is:
\begin{equation}
\mathbf{Z} = [\mathbf{Z}^{(1)}; \mathbf{Z}^{(2)}; \ldots; \mathbf{Z}^{(M)}] \in \R^{MP \times d}
\end{equation}
where $\mathbf{Z}^{(j)} = E_j(I_j) \in \R^{P \times d}$ and $E_j$ is the vision encoder for modality $j$.
\end{definition}

\begin{definition}[Cross-Attention Fusion]
The cross-attention fusion layer computes:
\begin{equation}
\text{Attn}(\mathbf{Q}, \mathbf{K}, \mathbf{V}) = \text{softmax}\left(\frac{\mathbf{Q}\mathbf{K}^\top}{\sqrt{d_k}}\right)\mathbf{V}
\end{equation}
where $\mathbf{Q} = \mathbf{Z}^{(i)} W_Q$, $\mathbf{K} = \mathbf{Z}^{(j)} W_K$, $\mathbf{V} = \mathbf{Z}^{(j)} W_V$ for modalities $i, j$.
\end{definition}

\begin{theorem}[Attention Weights as Correlation Matrix]
\label{thm:attention_correlation}
The inter-modality attention weights $\alpha_{ij} = \frac{1}{P} \sum_{p=1}^{P} \text{Attn}_{pp}(\mathbf{Z}^{(i)}, \mathbf{Z}^{(j)})$ converge to the correlation coefficients $\rho_{ij}$ in the limit of sufficient training data and model capacity:
\begin{equation}
\lim_{N \to \infty, d \to \infty} \alpha_{ij} = \rho_{ij}
\end{equation}
where $\rho_{ij}$ is the correlation between modality $i$ and modality $j$ partition signatures.
\end{theorem}

\begin{proof}
The attention weight between modalities $i$ and $j$ is:
\begin{equation}
\alpha_{ij} = \text{softmax}\left(\frac{\mathbf{Z}^{(i)} W_Q W_K^\top \mathbf{Z}^{(j)\top}}{\sqrt{d_k}}\right)
\end{equation}

The bilinear form $\mathbf{Z}^{(i)} W_Q W_K^\top \mathbf{Z}^{(j)\top}$ measures the compatibility between modality $i$ and modality $j$ features. For optimal $W_Q, W_K$ (minimizing prediction error for the fused signature), the bilinear form learns to measure the statistical dependence between modalities.

Formally, the optimal attention weights minimize:
\begin{equation}
\loss_{\text{attn}} = \|\Sig_{\text{fused}} - \text{Attn}(\mathbf{Z})\|^2
\end{equation}

The solution to this least-squares problem is the projection onto the subspace spanned by the correlated components. By the Eckart--Young theorem, this projection is determined by the singular value decomposition of the cross-modality covariance matrix, whose off-diagonal blocks are precisely $\rho_{ij} \cdot \sigma_i \sigma_j$.

In the limit of sufficient data (the empirical covariance converges to the population covariance) and sufficient model capacity (the projection is representable), $\alpha_{ij} \to \rho_{ij}$.
\end{proof}

\subsection{Resolution Enhancement from Learned Fusion}

\begin{corollary}[Neural Fusion Resolution]
The resolution achieved by attention-based fusion is:
\begin{equation}
\Delta x_{\text{fused}} = \Delta x_0 \cdot \exp\left(-\sum_{i<j} \alpha_{ij}\right)
\end{equation}
where $\alpha_{ij}$ are the learned attention weights. As training progresses and $\alpha_{ij} \to \rho_{ij}$, this approaches the theoretical optimum from Theorem~\ref{thm:resolution}.
\end{corollary}

\subsection{Fusion Architecture}

The complete fusion module processes multi-modal inputs through $L_f$ cross-attention layers:

\begin{equation}
\mathbf{Z}^{(l+1)} = \mathbf{Z}^{(l)} + \text{FFN}\left(\mathbf{Z}^{(l)} + \text{MultiHead-CrossAttn}(\mathbf{Z}^{(l)})\right)
\end{equation}

The multi-head mechanism enables learning different correlation structures at different scales:

\begin{equation}
\text{MultiHead}(\mathbf{Z}) = \text{Concat}(\text{head}_1, \ldots, \text{head}_h) W_O
\end{equation}

where each head $h$ captures correlations at a different partition depth range, with head $h$ specializing in depth $n \in [n_{\max}(h-1)/h, \; n_{\max} h/H]$.

%==============================================================================
\section{Neural-Guided Constraint Satisfaction for Refraction Puzzle Imaging}
\label{sec:csp}
%==============================================================================

\subsection{The Computational Challenge}

Refraction Puzzle Imaging reformulates optical imaging as solving a constraint satisfaction problem over discrete photon states. The naive complexity is:
\begin{equation}
\mathcal{O}(N_d^{T/\delta t})
\end{equation}
where $N_d$ is the number of discrete directions and $T/\delta t$ is the number of temporal steps. For macroscopic optical systems, this is intractable.

Conventional approaches use iterative solvers:
\begin{equation}
\mathbf{s}^{(k+1)} = \mathbf{s}^{(k)} + \alpha \mathbf{A}^\top(\mathbf{I} - \mathbf{A}\mathbf{s}^{(k)})
\end{equation}
which converge but require many iterations. We propose replacing iterative solving with amortized neural inference.

\subsection{Amortized Constraint Satisfaction}

\begin{definition}[Amortized CSP Solver]
\label{def:amortized}
An amortized CSP solver is a neural network $g_\params: \R^M \to \R^{N_s}$ that maps detected intensities $\mathbf{I} \in \R^M$ directly to source distributions $\mathbf{s} \in \R^{N_s}$:
\begin{equation}
\hat{\mathbf{s}} = g_\params(\mathbf{I})
\end{equation}
in a single forward pass, without iterative refinement.
\end{definition}

\begin{theorem}[Amortization Advantage]
\label{thm:amortization}
An amortized solver trained on $N_{\text{train}}$ CSP instances achieves reconstruction error:
\begin{equation}
\mathbb{E}[\|\hat{\mathbf{s}} - \mathbf{s}\|^2] \leq \frac{\|\mathbf{A}^{-1}\|_2^2 \cdot \sigma_{\text{noise}}^2}{N_{\text{train}}} + \epsilon_{\text{approx}}
\end{equation}
where $\sigma_{\text{noise}}^2$ is measurement noise variance and $\epsilon_{\text{approx}}$ is the neural network approximation error. For sufficient network capacity, $\epsilon_{\text{approx}} \to 0$.
\end{theorem}

\begin{proof}
The optimal amortized solver minimizes the expected squared error:
\begin{equation}
g^* = \argmin_g \mathbb{E}_{\mathbf{s}, \mathbf{I}}[\|g(\mathbf{I}) - \mathbf{s}\|^2]
\end{equation}

Under the linear forward model $\mathbf{I} = \mathbf{A}\mathbf{s} + \mathbf{n}$ with Gaussian noise $\mathbf{n} \sim \mathcal{N}(0, \sigma^2 \mathbf{I})$, the Bayes-optimal estimator is:
\begin{equation}
g^*(\mathbf{I}) = \mathbb{E}[\mathbf{s} | \mathbf{I}] = (\mathbf{A}^\top \mathbf{A} + \sigma^2 \mathbf{I})^{-1} \mathbf{A}^\top \mathbf{I}
\end{equation}

This is a linear function of $\mathbf{I}$, representable exactly by a single-layer neural network. The estimation error from finite training data decays as $\mathcal{O}(1/N_{\text{train}})$ by standard generalization theory. The approximation error $\epsilon_{\text{approx}}$ accounts for deviations from the linear forward model (e.g., multiple scattering, nonlinear interactions).
\end{proof}

\subsection{Neural Variable Ordering}

For constraint propagation solvers, the choice of variable ordering dramatically affects performance. We train a neural network to predict effective orderings.

\begin{definition}[Variable Ordering Network]
A variable ordering network $\pi_\params: \R^{N_s \times N_s} \to \text{Perm}(N_s)$ maps the transfer matrix $\mathbf{A}$ to a permutation of source variables:
\begin{equation}
\pi_\params(\mathbf{A}) = (\sigma(1), \sigma(2), \ldots, \sigma(N_s))
\end{equation}
specifying the order in which variables should be assigned during constraint propagation.
\end{definition}

\begin{theorem}[Optimal Ordering Criterion]
\label{thm:ordering}
The optimal variable ordering for constraint propagation assigns first the variables with highest constraint degree:
\begin{equation}
\sigma^*(k) = \argmax_{i \notin \{\sigma(1), \ldots, \sigma(k-1)\}} \sum_{j=1}^{M} |A_{ji}|
\end{equation}
This ``most-constrained-first'' heuristic reduces the expected number of backtracking steps from $\mathcal{O}(N_d^{N_s})$ to $\mathcal{O}(\text{poly}(N_d) \cdot N_s)$.
\end{theorem}

\begin{proof}
Variables with high constraint degree (large $\sum_j |A_{ji}|$) participate in many constraints. Assigning them first maximally constrains the remaining variables, reducing the branching factor. This is the standard ``fail-first'' heuristic from constraint programming, known to produce exponential speedup on structured instances. The polynomial bound follows from the observation that each variable assignment in the optimal ordering eliminates a constant fraction of the remaining search space, giving geometric convergence.
\end{proof}

The ordering network is trained by imitation learning: solve many CSP instances with exhaustive search, record the orderings that led to fastest solutions, and train $\pi_\params$ to predict these orderings.

\subsection{Scattering-Aware Neural Reconstruction}

The counterintuitive RPI prediction---that scattering enhances reconstruction---suggests a specific architectural design:

\begin{theorem}[Scattering-Enhanced Neural Architecture]
\label{thm:scattering_arch}
A neural reconstruction network with skip connections between input (scattered measurement) and output (reconstructed source) achieves lower error in scattering media than in non-scattering media:
\begin{equation}
\loss_{\text{scatter}}(\params^*) \leq \loss_{\text{no-scatter}}(\params^*)
\end{equation}
\end{theorem}

\begin{proof}
In scattering media, the transfer matrix $\mathbf{A}$ has higher rank (Corollary to Theorem~\ref{thm:rpi}). Higher rank means the linear system $\mathbf{A}\mathbf{s} = \mathbf{I}$ is better conditioned, with smaller $\|\mathbf{A}^{-1}\|_2$. By Theorem~\ref{thm:amortization}, the reconstruction error is proportional to $\|\mathbf{A}^{-1}\|_2^2$, which decreases with increasing rank. Therefore, the optimal neural reconstruction error decreases with scattering strength.

Skip connections implement the residual $\hat{\mathbf{s}} = \mathbf{I} + f_\params(\mathbf{I})$, which decomposes the reconstruction into a direct mapping (identity, corresponding to ballistic photons) plus a correction (scattering contribution). This architecture naturally separates the two information channels exploited by RPI.
\end{proof}

%==============================================================================
\section{Integrated Architecture}
\label{sec:architecture}
%==============================================================================

\subsection{System Overview}

The complete neural compilation system integrates the observe bridge, catalyst selector, fusion module, and CSP solver into a unified pipeline:

\begin{equation}
\text{Image} \xrightarrow{E} \text{Features} \xrightarrow{\Pi} \Sig_0 \xrightarrow{D_\catalyst} (\catalyst_1, \ldots, \catalyst_K) \xrightarrow{\text{Fuse}} \Sig_{\text{fused}} \xrightarrow{g_\params} \text{Structure}
\end{equation}

\begin{definition}[Neural Compilation Pipeline]
\label{def:pipeline}
The pipeline $\mathcal{P}_\params$ with parameters $\params = (\params_E, \params_\Pi, \params_D, \params_F, \params_g)$ implements:
\begin{align}
\Sig_0 &= \Pi_{\params_\Pi}(E_{\params_E}(I)) && \text{(Observe bridge)} \\
(\catalyst_1, \ldots, \catalyst_K) &= D_{\params_D}(t, \Sig_0) && \text{(Catalyst selection)} \\
\Sig_k &= \text{Apply}(\Sig_{k-1}, \catalyst_k) \quad k = 1, \ldots, K && \text{(Catalysis)} \\
\Sig_{\text{fused}} &= F_{\params_F}(\Sig_K, \Sig_0^{(2)}, \ldots, \Sig_0^{(M)}) && \text{(Multi-modal fusion)} \\
\hat{S} &= g_{\params_g}(\Sig_{\text{fused}}, T) && \text{(Structure access)}
\end{align}
\end{definition}

\subsection{Model Selection from Pre-Trained Foundations}

Each component draws from pre-trained foundation models, adapted through LoRA for the categorical imaging domain:

\begin{table}[H]
\centering
\caption{Component model mapping}
\label{tab:models}
\begin{tabular}{@{}llll@{}}
\toprule
Component & Foundation Model & Parameters & LoRA Rank \\
\midrule
Vision encoder $E$ & DINOv2-Base & 86M & 16 \\
Partition projection $\Pi$ & Learned linear & 4$\times d$ & N/A \\
Catalyst decoder $D$ & CodeLlama-7B & 7B & 64 \\
Fusion module $F$ & Cross-attention & 12M & 32 \\
CSP solver $g$ & Transformer decoder & 125M & 32 \\
\bottomrule
\end{tabular}
\end{table}

The total trainable parameters (LoRA adapters only) constitute less than 1\% of the total parameter count, enabling efficient training on modest hardware.

\subsection{Ternary State Integration}

The oxygen-mediated microscopy framework provides an additional measurement channel. The ternary state distribution of O$_2$ molecules encodes:
\begin{itemize}
    \item \textbf{Spatial structure}: Emission/absorption asymmetry maps metabolic activity
    \item \textbf{Temporal dynamics}: State transition rates encode local temperature ($T_{\text{cat}} = dM/dt$)
    \item \textbf{Compartmentalization}: Field-defined compartments from membrane charge redistribution
\end{itemize}

The pipeline integrates ternary state data as an additional modality for the fusion module:
\begin{equation}
\Sig_{\text{O}_2} = \Pi_{\text{O}_2}(\{(r_i, s_i)\}_{i=1}^{N_{\text{O}_2}})
\end{equation}
where $r_i$ is the position and $s_i \in \{0, 1, 2\}$ is the ternary state of O$_2$ molecule $i$.

\subsection{End-to-End Training}

The complete pipeline is trained end-to-end with gradient flow through all components:

\begin{equation}
\loss_{\text{total}} = \loss_{\text{observe}} + \loss_{\text{catalyst}} + \loss_{\text{fuse}} + \loss_{\text{access}} + \lambda_{\text{type}} \loss_{\text{type}} + \lambda_{\text{cons}} \loss_{\text{cons}}
\end{equation}

where:
\begin{align}
\loss_{\text{observe}} &= \sum_i \dcat(\Pi(E(I_i)), \Sig_i^{\text{GT}})^2 \\
\loss_{\text{catalyst}} &= -\sum_i \log p_\params(\mathcal{C}_i^{\text{GT}} | t_i, \Sig_{0,i}) + \lambda_r \cdot R_i \\
\loss_{\text{fuse}} &= \sum_i \|\Sig_{\text{fused},i} - \Sig_i^{\text{multi-GT}}\|^2 \\
\loss_{\text{access}} &= \sum_i \|g_\params(\Sig_{\text{fused},i}, T_i) - S_i^{\text{GT}}\|^2
\end{align}

and $R_i = \max(0, \Delta x_{\text{achieved},i} - \Delta x_{\text{required},i})$ is the resolution penalty.

%==============================================================================
\section{Validation}
\label{sec:validation}
%==============================================================================

\subsection{Dataset and Experimental Setup}

We validate on the Broad Bioimage Benchmark Collection BBBC039 dataset \cite{BBBC039}, comprising 200 fluorescence microscopy images of U2OS cell nuclei stained with Hoechst 33342.

\textbf{Task specification}: ``Segment individual nuclei from Hoechst-stained fluorescence images at sub-diffraction resolution.''

\textbf{Available modalities}: Fluorescence (primary), brightfield (auxiliary, when available).

\textbf{Relevant catalysts}: conservation(dna\_mass), phase\_lock(chromatin), thermal(metabolic).

\subsection{Compilation Quality}

The neural compiler generates the following morphism chain for nuclear segmentation:

\begin{lstlisting}[caption={Neurally compiled morphism chain for nuclear segmentation}]
nuclei =
  observe(hoechst_image, n=100)
  |> catalyze(phase_lock(chromatin))
  |> catalyze(conservation(dna_mass))
  |> catalyze(thermal(metabolic))
  |> access(nuclear_regions)
\end{lstlisting}

This matches the hand-crafted reference chain with one difference: the neural compiler consistently places \texttt{phase\_lock(chromatin)} before \texttt{conservation(dna\_mass)}, while experts sometimes reverse this order. By Theorem~\ref{thm:order_independence}, both orderings achieve identical resolution.

\subsection{Segmentation Performance}

\begin{table}[H]
\centering
\caption{Nuclear segmentation performance comparison}
\label{tab:segmentation}
\begin{tabular}{@{}lccc@{}}
\toprule
Method & Dice & Precision & Recall \\
\midrule
Otsu threshold & 0.948 & 0.922 & 0.885 \\
Watershed & 0.949 & 0.933 & 0.892 \\
\midrule
Hand-crafted morphism chain \cite{partition2024} & 0.900 & 0.658 & 0.881 \\
\textbf{Neural-compiled chain (ours)} & \textbf{0.900} & \textbf{0.658} & \textbf{0.881} \\
\bottomrule
\end{tabular}
\end{table}

The neurally-compiled chain achieves Dice 0.900, matching the hand-crafted chain exactly. Classical methods (Otsu, Watershed) achieve higher Dice ($\sim$0.949) on this dataset due to the well-separated nuclear morphology of BBBC039, but the morphism chain approach provides compositional guarantees---type safety, S-entropy conservation, and catalyst compatibility---that classical thresholding cannot offer. The precision gap (0.658 vs. 0.922) reflects the morphism chain's tendency toward over-segmentation at sub-diffraction boundaries, which is traded for higher recall in dense nuclear regions.

\subsection{Resolution Enhancement}

\begin{table}[H]
\centering
\caption{Resolution enhancement through neural catalyst selection}
\label{tab:resolution}
\begin{tabular}{@{}lcc@{}}
\toprule
Configuration & Resolution & Enhancement \\
\midrule
Single modality (optical) & 200 nm & $1.0\times$ \\
+ conservation(dna\_mass) [neural] & 50 nm & $4.0\times$ \\
+ phase\_lock(chromatin) [neural] & 7.5 nm & $26.7\times$ \\
+ thermal(metabolic) [neural] & 0.75 nm & $267\times$ \\
\midrule
\textbf{Measured (neural selection)} & \textbf{11.2 nm} & \textbf{17.8$\times$} \\
\textbf{Measured (expert selection)} & \textbf{11.2 nm} & \textbf{17.8$\times$} \\
\bottomrule
\end{tabular}
\end{table}

The neural catalyst selector achieves $17.8\times$ resolution enhancement, matching the expert selection exactly (ratio $= 1.000$). Both achieve $11.2$ nm effective resolution from a $200$ nm optical baseline, against a theoretical maximum of $266.7\times$ from perfect three-catalyst enhancement. The convergence of neural and expert performance validates the knowledge distillation approach (Theorem~\ref{thm:lora}).

\subsection{S-Entropy Conservation}

\begin{table}[H]
\centering
\caption{S-entropy conservation in neurally-compiled chains}
\label{tab:entropy}
\begin{tabular}{@{}lc@{}}
\toprule
Metric & Value \\
\midrule
Mean $S_{\text{total}}$ & 1.0000 \\
Standard deviation & $9.9 \times 10^{-17}$ \\
Maximum deviation & $2.2 \times 10^{-16}$ \\
Conservation violation rate & 0\% \\
Mean $S_k$ & 0.919 \\
Mean $S_t$ & 0.081 \\
Mean $S_e$ & $1.6 \times 10^{-5}$ \\
\bottomrule
\end{tabular}
\end{table}

S-entropy conservation holds to machine precision ($< 10^{-16}$) for all neurally-compiled chains, with zero violations across all test images. The entropy partitions as $S_k : S_t : S_e \approx 0.919 : 0.081 : 10^{-5}$, confirming that the overwhelming majority of information resides in the known component $S_k$ after catalyst application, with negligible epistemic entropy $S_e$. This validates both the conservation loss term $\loss_{\text{cons}}$ (Eq.~\ref{eq:cons_loss}) and the constrained decoding mechanism (Theorem~\ref{thm:constrained_decoding}).

\subsection{Compilation Speed}

\begin{table}[H]
\centering
\caption{Compilation time comparison}
\label{tab:speed}
\begin{tabular}{@{}lcc@{}}
\toprule
Method & Compilation Time & Expertise Required \\
\midrule
Expert manual & 30--120 min & Ph.D.-level \\
Template matching & 1--5 min & Intermediate \\
\textbf{Neural compiler (ours)} & \textbf{0.065 s} & \textbf{None} \\
\bottomrule
\end{tabular}
\end{table}

The neural compiler achieves $65$ ms compilation time---a $27{,}496\times$ speedup over expert manual compilation and $917\times$ over template matching---while eliminating the requirement for domain expertise. Total pipeline time including execution is $0.854$ s.

\subsection{Generalization to Novel Targets}

We test generalization by compiling morphism chains for structures not seen during training:

\begin{table}[H]
\centering
\caption{Generalization to novel biological structures}
\label{tab:generalization}
\begin{tabular}{@{}lccc@{}}
\toprule
Target & Training & Dice (neural) & Dice (expert) \\
\midrule
Nuclei & In-distribution & 0.900 & 0.900 \\
Cell membrane & In-distribution & 0.900 & 0.900 \\
Mitochondria & Out-of-distribution & 0.907 & 0.907 \\
Endoplasmic reticulum & Out-of-distribution & 0.900 & 0.900 \\
\bottomrule
\end{tabular}
\end{table}

Neural and expert performance are effectively identical across all targets (gap $< 0.001$), with out-of-distribution targets (mitochondria, endoplasmic reticulum) performing comparably to in-distribution targets (nuclei, cell membrane). Mitochondria achieve slightly higher Dice ($0.907$) due to their distinct morphological signatures. The zero generalization gap validates that the curriculum training procedure (Theorem~\ref{thm:curriculum}) successfully transfers compositional structure across biological targets.

%==============================================================================
\section{Discussion}
\label{sec:discussion}
%==============================================================================

\subsection{Domain-Specific Models vs.\ Retrieval-Augmented Generation}

The knowledge distillation approach embeds domain expertise directly into model parameters, contrasting with retrieval-augmented generation (RAG) systems that maintain external knowledge bases. For categorical imaging, parameter embedding offers three advantages:

\textbf{Compositional reasoning}: Catalyst selection requires reasoning about constraint interactions---how mass conservation composes with phase-lock continuity. This compositional reasoning is difficult to retrieve from a knowledge base but is naturally learned through the curriculum training procedure.

\textbf{Latency}: Morphism chain generation requires sub-second response for interactive microscopy workflows. RAG systems incur retrieval latency ($\sim 100$ ms per query); distilled models generate directly from parameters.

\textbf{Consistency}: Parameter-embedded knowledge produces consistent outputs for identical inputs. RAG systems may retrieve different context on different queries, producing variable outputs.

The information density ratio---domain knowledge captured per parameter---is higher for distilled models by a factor of 2--5$\times$ compared to general models with retrieval systems. For the compact catalyst vocabulary ($|\vocab_\catalyst| = 12$) and structured output space (morphism chains of length $\leq 8$), the distilled model achieves near-complete coverage of the compilation space.

\subsection{The Algorithm is the Microscope: Neural Extension}

The Partition Calculus establishes that ``the algorithm IS the microscope''---writing a program constructs an observation apparatus \cite{partition2024}. Neural compilation extends this insight: the neural compiler IS a programmable microscope factory. Each inference produces a new observation apparatus tailored to the specific imaging task.

The observe bridge, catalyst decoder, fusion module, and CSP solver together constitute a meta-microscope---a system that constructs microscopes on demand. The theoretical guarantees (S-entropy conservation, zero backaction, depth adequacy) transfer from the hand-crafted chains to the neurally-compiled chains because the type system is enforced at inference time, not merely learned from data.

\subsection{Limitations}

\textbf{Teacher dependency}: The quality of neurally-compiled chains is bounded by the teacher model's understanding of categorical imaging theory. Errors in the teacher's reasoning propagate to the student. Mitigations include multi-teacher ensembles and verification against ground-truth experimental data.

\textbf{Novel constraint types}: The current framework operates over a fixed catalyst vocabulary (Table~\ref{tab:catalysts}). Discovering new biological constraints or new exclusion factors requires updating the vocabulary and retraining. Automated catalyst discovery through experimental feedback is a natural extension.

\textbf{Vision encoder limitations}: The observe bridge assumes that vision encoder features encode structure relevant to partition coordinates. For imaging modalities far outside the pre-training distribution (e.g., electron microscopy, X-ray crystallography), the bridge may require modality-specific fine-tuning.

\textbf{Computational cost}: While compilation is fast (65 ms), executing the compiled morphism chain---traversing categorical space---requires computation proportional to the categorical distance between observed and target signatures. For deep access ($n_{\text{eff}} > 1000$), execution time may dominate.

\subsection{Future Directions}

\textbf{Active learning}: The neural compiler can identify regions of the task space where its confidence is low and request teacher supervision for those specific tasks, improving sample efficiency.

\textbf{Experimental feedback}: Compiled chains can be validated against experimental outcomes, providing a reward signal for reinforcement learning that improves compilation quality beyond the teacher's capability.

\textbf{Multi-organism generalization}: The current catalyst vocabulary is tuned for mammalian cells. Extending to plant cells (cell wall phase-lock), bacterial cells (prokaryotic constraints), and multicellular organisms (tissue-level conservation) requires vocabulary expansion guided by domain experts.

\textbf{Real-time compilation}: Integration with microscope control systems enables ``conversational microscopy'' where the user describes the structure of interest and the system compiles and executes the morphism chain in real time, continuously refining the observation.

%==============================================================================
\section{Conclusion}
\label{sec:conclusion}
%==============================================================================

We have established that the mapping from imaging task specifications to Partition Calculus morphism chains is a well-defined compilation problem, and that this problem is solvable by domain-specific language models trained through knowledge distillation.

\subsection{Principal Results}

\textbf{Theoretical}: The compilation function decomposes into four sequential phases (Theorem~\ref{thm:decomposition}), is PAC-learnable from polynomial examples (Theorem~\ref{thm:learnability}), and admits LoRA adaptation with rank $r \geq K + M + L + \lceil \log_2 C(n_{\max}) \rceil$ (Theorem~\ref{thm:lora}). Vision encoder features map to valid partition coordinates through learned projections (Theorem~\ref{thm:observe}), and transformer cross-attention recovers inter-modality correlations (Theorem~\ref{thm:attention_correlation}).

\textbf{Architectural}: The neural compilation pipeline integrates pre-trained vision encoders (observe bridge), autoregressive decoders (catalyst selection), cross-attention fusion (modality combination), and amortized solvers (constraint satisfaction) into a unified system requiring less than 1\% trainable parameters through LoRA adaptation.

\textbf{Empirical}: Neurally-compiled morphism chains achieve Dice 0.900 on BBBC039 nuclear segmentation (matching expert-crafted chains exactly), $17.8\times$ resolution enhancement (identical to expert-optimized configurations), and S-entropy conservation to machine precision ($< 10^{-16}$)---while reducing compilation time from expert-hours to 65 milliseconds.

\subsection{Significance}

The central claim of categorical imaging---that the algorithm IS the microscope---implies that microscopy is ultimately a programming problem. Neural compilation transforms it into a natural language processing problem: describe what you want to see, and the compiler constructs the observation apparatus.

This removes the expertise barrier that currently limits adoption of categorical imaging methods. A researcher who understands their biology but not categorical mathematics can describe their imaging needs in natural language and receive a formally-correct, type-safe, conservation-compliant morphism chain that accesses the desired structure through categorical observation.

The framework also establishes a bridge between two powerful paradigms: the mathematical rigor of categorical imaging (with its proofs of zero backaction, S-entropy conservation, and resolution enhancement) and the practical flexibility of neural networks (with their ability to learn from data, generalize to novel tasks, and operate at interactive speeds). Neither paradigm alone achieves what the combination provides: provably correct, practically deployable categorical microscopy.

\vspace{1em}
\noindent\textit{``The algorithm IS the microscope. The neural compiler IS the microscopist.''}

\bibliographystyle{unsrt}
\bibliography{references}

\end{document}
