Traditional video playback exhibits temporal symmetry: scrubbing forward plays the video forward in time, while scrubbing backward reverses temporal direction. This symmetry, while intuitive, is not fundamental—it arises from indexing frames by chronological time rather than by information-theoretic properties. We present a novel video representation framework where frames are organized in \textit{categorical entropy coordinates} rather than temporal coordinates, yielding \textit{temporally irreversible playback}: regardless of scrubbing direction, the video always depicts forward temporal evolution. 

The key innovation is the introduction of \textit{Motion Picture Maxwell Demons}—categorical observers that enforce thermodynamic constraints on video traversal. Each frame is assigned cumulative entropy coordinates $(S_k, S_t, S_e, S_{\text{cum}})$ computed from pixel intensity distributions, optical flow, and temporal correlations. Since entropy production is monotonically increasing, backward scrubbing cannot decrease cumulative entropy $S_{\text{cum}}$; instead, the system accesses alternative forward-evolving paths through a dual-membrane temporal structure.

We establish the mathematical foundations through: (1) temporal S-entropy coordinate systems extending Shannon entropy to video sequences, (2) gas molecular dynamics models treating pixel motion as Brownian trajectories with collective behavior, (3) dual-membrane frame representations providing conjugate forward-time paths analogous to amplitude-phase complementarity, and (4) temporal Maxwell demons that validate frame transitions against thermodynamic consistency.

Validation on biological video sequences (cell migration, embryonic development, fluid dynamics) demonstrates successful irreversible playback with $>95\%$ visual continuity while enforcing strict entropy monotonicity. The framework enables novel applications including tamper-evident surveillance video, irreversible scientific visualization, and pedagogical tools where temporal reversal is physically meaningless.

This work unifies information theory, thermodynamics, and video representation, suggesting that entropy coordinates may be more fundamental than chronological time for organizing sequences of irreversible processes. The approach generalizes to any sequential data exhibiting monotonic information production, establishing categorical entropy as a universal coordinate system for temporal data.

\textbf{Keywords:} Video representation, entropy coordinates, Maxwell demon, temporal irreversibility, information theory, thermodynamics, categorical computation

